Galaxy Zoo: Morphological Classifications for 150,000 Galaxies in HST Legacy Imaging

### 2016-05-26 writeup/gz_hubble_data.tex

The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the image as seen on screen.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. The combination of all
such parameters forms a high-dimensional space, and it is not clear how to
separate this into individual effects. Instead, the method used here employs
only two parameters as a simple model of the effect of redshift brightness on classification: the
$r^\prime$-band surface brightness ($\mu_r$; intrinsic) and redshift ($z$;
extrinsic).

%This dependence of the vote fraction on $\mu_r$ and \zsim{} is measured using the \ferengi{} images over all the evolutionary correction factors. 
We model the debiased vote fraction $f_{\mu,z}$ for an
artificially-redshifted galaxy as:

@@

\noindent where $f_{\mu,z=0.3}$ is the vote fraction at the
lowest redshift in the \ferengi{} bins ($z_0=0.3$). Here, $\zeta$ is a model
parameter that controls the rate at which \ffeatures{} decreases with increasing
redshift (and that may depend on other galaxy properties). This formula fits
the data relatively well (with very few exceptions, the vote fractions for
featured galaxies decrease monotonically with increasing redshift), and the
exponential function bounds the observed vote fractions between $f_{\mu,z=0.3}$
and zero. Figure~\ref{fig:zeta_examples} shows the change in vote fraction and
the best-fit model for a random sample of galaxies.

@@

The data from Figure~\ref{fig:zeta_examples} show that there is a clear
dependence on redshift for the observed changes in \ffeatures. To examine
whether a better fit should involve different parameters, we examine the global
dependence of $\zeta$ on the galaxy surface brightness. This is motivated by
the fact that brighter galaxies would presumably be easier to identify;
\citet{bam09} and \citet{wil13} found changes in \ffeatures{} depending on a
combination of physical size and absolute magnitude. To simplify our model, we
subsume these into a single parameter $\mu$.

@@

Figure~\ref{fig:f_vs_f} shows the change in \ffeatures{} for the \ferengi{}
galaxies relative to their lowest simulated redshift. This analysis only uses
galaxies whose lowest simulated redshift image was $z_\mathrm{sim}=0.3$ and
with detectable surface brightness measurements from \sextractor; this includes
3,950 of the total 6,466~images originally processed. For each simulated
redshift value \zsim{} at a fixed surface brightness $\mu$, \ffeaturesz, the
value measured at that simulated redshift, is plotted against \ffeaturesrest,
the value measured for the same galaxy at \zsim$=0.3$.

@@

The criteria for determining whether a region of this space is single-valued,
and therefore correctable, is as follows: In each surface brightness and
redshift bin, the relationship between \ffeaturesz{} and \ffeaturesrest{} is
modeled by fitting the data with polynomials of degrees $n=3$, 2, and 1, and
using the best formal fit out of the three as measured by the sum of the
residuals. These fits are shown as the dashed black lines in
Figure~\ref{fig:f_vs_f}. Flat regions of the bins are areas in which there is
not a clear single-valued relationship between \ffeaturesz{} and
\ffeaturesrest. We quantify this by measuring the slope of the best-fit
polynomical; regions of the bins with a slope less than 0.4 are
considered \emph{not} one-to-one, and therefore \ffeaturesz{} cannot be boosted
to its \ffeaturesrest{} value.  Galaxies in this region are referred to as the
\emph{lower limit} sample, because the most stringent correction available is
that the weighted \ffeatures{} is a lower limit to the true value.  These
regions are highlighted in blue in Figure~\ref{fig:f_vs_f}. Uncolored (white)
regions of the plot have sufficiently high slopes to consider the relationship
as single-valued; galaxies in these regions are considered ``correctable'', and
only these are used in measuring the parameters for the $\zeta$ function
(Section~\ref{ssec:zeta}). Only surface brightness/redshift bins with at least
5~galaxies were considered; regions with fewer than 5~galaxies are considered
to have ``not enough information'' to determine the \ffeaturesz{} and
\ffeaturesrest{} relationship, colored gray in Figure~\ref{fig:f_vs_f}.

>>>>

The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the galaxy.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. The combination of all
such parameters forms a high-dimensional space, and it is not clear how to
separate this into individual effects. Instead, the method used here employs
only two parameters as a simple model of the effect of redshift brightness on classification: the
$i^\prime$-band surface brightness ($\mu_\mathrm{i}$; intrinsic) and redshift ($z$;
extrinsic).

We model the debiased vote fraction $f_{\mu,z}$ for an
artificially-redshifted galaxy as:

@@

\noindent where $f_{\mu,z=0.3}$ is the vote fraction at the lowest redshift in
the artificially-redshifted sample ($z_0=0.3$). Here, $\zeta$ is a parameter that
controls the rate at which \ffeatures{} decreases with increasing redshift (and
that may depend on other galaxy properties). This function bounds the observed
vote fractions between $f_{\mu,z=0.3}$ and zero based on two assumptions: 1)
the vote fractions for featured galaxies decrease monotonically with increasing
redshift, therefore restricting vote fractions for a given galaxy to be less
than $f_{\mu,z=0.3}$ (which is almost always true of the data), and 2) the
vote fractions cannot be less than zero (which is always true).

@@

Figure~\ref{fig:zeta_examples} shows the change in vote fraction and the
best-fit model for a random set of galaxies in the \ferengi{} sample. The
results show that there is a clear dependence on redshift for the observed
changes in \ffeatures. To examine whether a better fit should involve
additional parameters, we examine the global dependence of $\zeta$ on the
galaxy surface brightness. This is motivated by the fact that brighter galaxies
would presumably be easier to identify; \citet{bam09} and \citet{wil13} found
changes in \ffeatures{} depending on a combination of physical size and
absolute magnitude. To simplify our model, we subsume these into a single
parameter $\mu$.

@@

Figure~\ref{fig:f_vs_f} shows the change in \ffeatures{} for the \ferengi{}
galaxies relative to their lowest simulated redshift. This analysis only uses
galaxies whose lowest simulated redshift image was $z_\mathrm{sim}=0.3$ and
with detectable surface brightness measurements from \sextractor{} \citep{mel16}; this includes
3,950 of the total 6,466~images originally processed. For each simulated
redshift value \zsim{} at a fixed surface brightness $\mu$, \ffeaturesz, the
value measured at that simulated redshift, is plotted against \ffeaturesrest,
the value measured for the same galaxy at \zsim$=0.3$.

@@

The criteria for determining whether a region of this space is single-valued,
and therefore correctable, is as follows: In each surface brightness and
redshift bin, the relationship between \ffeaturesz{} and \ffeaturesrest{} is
modeled by fitting the data with polynomials of degrees $n=3$, 2, and 1, and
using the best formal fit out of the three as measured by the sum of the
residuals. These fits are shown as the dashed black lines in
Figure~\ref{fig:f_vs_f}. Flat regions of the bins are areas in which there is
not a clear single-valued relationship between \ffeaturesz{} and
\ffeaturesrest. We quantify this by measuring the slope of the best-fit
polynomical; regions of the bins with a slope less than 0.4 are
considered \emph{not} one-to-one, and therefore \ffeaturesz{} cannot be boosted
to its \ffeaturesrest{} value.  Galaxies in this region are referred to as the
\emph{lower limit} sample, because the most stringent correction available is
that the weighted \ffeatures{} is a lower limit to the true value.  These
regions are highlighted in blue in Figure~\ref{fig:f_vs_f}. Uncolored (white)
regions of the plot have sufficiently high slopes to consider the relationship
as single-valued; galaxies in these regions are considered ``correctable'', and
only these are used in measuring the parameters for the $\zeta$ function
(Section~\ref{ssec:zeta}). Only surface brightness/redshift bins with at least
5~galaxies were considered; regions with fewer than 5~galaxies are considered
to have ``not enough information'' to determine the \ffeaturesz{} and
\ffeaturesrest{} relationship, colored gray in Figure~\ref{fig:f_vs_f}. These galaxies are hereafter referred to as the ``NEI'' sample.
